KidsAIR initial moisture content data work

### 2020-06-24 child_health_data_cleaning.Rmd
check home char across winters

nn_birthinfancy_initial <- read_csv(paste0(file_path, "Input/NN/birthinfancy.csv"), 
                                  na = c("NULL", "")) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         child_id_num = as.character(ChildID),
         area = "NN",
         gender = as.factor(Gender),
         bw_lbs = as.numeric(BirthweightLbs),
         bw_oz = as.numeric(BirthWeightOz),
         birth_weight_oz = as.numeric((bw_lbs*16)+bw_oz),
         mom_age = as.numeric(AgeOfMom),
         breastfeed_current = as.factor(CurrentlyBreastfeeding),
         breastfeed_past = as.factor(DidBreastfeed),
         bf_years = as.numeric(HowLongYears),
         bf_months = as.numeric(HowLongMonths),
         breastfeed_months = as.numeric((bf_years*12)+bf_months),
         flu_shot = as.factor(FluShot),
         flu_shot_planned = as.factor(FluShotPlanned),
         asthma_meds = as.factor(AsthmaMeds),
         child_school = as.factor(Attend),
         school_hours = as.numeric(HoursDay),
         school_days = as.numeric(DaysWeek),
         school_hours_week = as.numeric(school_hours*school_days),
         sibling_school = as.factor(SibSchool)) %>% 
  replace_with_na(replace = list(bw_lbs = 0, birth_weight_oz = 0, mom_age = 0)) %>%
  select(home_winter_id:sibling_school)

>>>>

nn_birthinfancy_initial <- read_csv(paste0(file_path, "Input/NN/birthinfancy.csv"), 
                                  na = c("NULL", "")) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         child_id_num = as.character(ChildID),
         area = "NN",
         gender = as.factor(Gender),
         bw_lbs = as.numeric(BirthweightLbs),
         bw_oz = as.numeric(BirthWeightOz),
         birth_weight_oz = as.numeric((bw_lbs*16)+bw_oz),
         mom_age = as.numeric(AgeOfMom),
         breastfeed_current = as.factor(CurrentlyBreastfeeding),
         breastfeed_past = as.factor(DidBreastfeed),
         bf_years = as.numeric(HowLongYears),
         bf_months = as.numeric(HowLongMonths),
         breastfeed_months = as.numeric((bf_years*12)+bf_months),
         flu_shot = as.factor(FluShot),
         flu_shot_planned = as.factor(FluShotPlanned),
         asthma_meds = as.factor(AsthmaMeds),
         child_school = as.factor(Attend),
         school_hours = as.numeric(HoursDay),
         school_days = as.numeric(DaysWeek),
         school_hours_week = as.numeric(school_hours*school_days),
         sibling_school = as.factor(SibSchool)) %>% 
  replace_with_na(replace = list(bw_lbs = 0, birth_weight_oz = 0, mom_age = 0)) %>%
  # checked young ages with field data - correcting here
  mutate(mom_age = if_else(mom_age == 14, 20, mom_age),
         mom_age = if_else(mom_age == 10 | mom_age == 12, 21, mom_age)) %>% 
  select(home_winter_id:sibling_school)

### 2020-06-24 demographics_summary.Rmd
check home char across winters

kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids.rds"))
demographics_initial <- read_rds(paste0(file_path, "Output/demographics_clean.rds")) 
pm_clean <- read_rds(paste0(file_path, "Output/pm_clean.rds"))
```

>>>>

kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds"))
demographics <- read_rds(paste0(file_path, "Output/demographics_final.rds"))
pm_clean <- read_rds(paste0(file_path, "Output/pm_clean.rds"))
```

### 2020-06-24 home_char_act_summary.Rmd
check home char across winters

---
title: "Kids home characteristics/activity summary"
author: "Ethan Walker"
date: "Started 15 Jan 2020, Updated 25 March 2020"
output:
  pdf_document: default
  powerpoint_presentation: default
---

@@

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

@@

>>>>

---
title: "Kids home characteristics/activity summary"
author: "Ethan Walker"
date: "Started 15 Jan 2020, Updated 23 June 2020"
output:
  pdf_document: default
  powerpoint_presentation: default
---

@@

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
library(forcats)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

@@

# Check Winter 1 vs Winter 2 data for home characteristics
```{r}
home_char_check_num <- home_char_clean %>% 
  group_by(home) %>% 
  arrange(home, winter_id) %>% 
  pivot_longer(cols = c(residents_total:residents_under18, home_year_built:home_furry), 
               names_to = "var", values_to = "value") %>% 
  group_by(home, var) %>% 
  arrange(home, var, winter_id) %>% 
  mutate(data_check = if_else(value == lead(value), 1, 0),
         data_check = if_else(is.na(data_check), lag(data_check), data_check)) %>% 
  filter(data_check == 0) %>% 
  select(area, home, winter_id, var, value) %>% 
  mutate(value = as.numeric(value),
         data_check = if_else(var == "residents_total" & abs(value - lead(value)) > 1, 1, 0),
         data_check = if_else(var == "residents_under5" & abs(value - lead(value)) > 1, 1, data_check),
         data_check = if_else(var == "residents_under18" & abs(value - lead(value)) > 1, 1, data_check),
         data_check = if_else(var == "home_year_built" & abs(value - lead(value)) > 10, 1, data_check),
         data_check = if_else(var == "home_sqft" & abs(value - lead(value)) > 200, 1, data_check),
         data_check = if_else(var == "home_floors" & abs(value - lead(value)) > 1, 1, data_check),
         data_check = if_else(var == "home_bedrooms" & abs(value - lead(value)) > 1, 1, data_check),
         data_check = if_else(var == "home_windows" & abs(value - lead(value)) > 5, 1, data_check),
         data_check = if_else(var == "home_pets" & abs(value - lead(value)) > 2, 1, data_check),
         data_check = if_else(var == "home_furry" & abs(value - lead(value)) > 2, 1, data_check),
         data_check = if_else(is.na(data_check), lag(data_check), data_check),
         value = as.character(value)) %>% 
  filter(data_check == 1) %>% 
  select(area, home, winter_id, var, value)

write_csv(home_char_check_num, paste0(file_path, "Output/home_char_check_num.csv"))

home_char_check_fact <- home_char_clean %>% 
  group_by(home) %>% 
  arrange(home, winter_id) %>% 
  pivot_longer(cols = c(home_type, stove_age, stove_cert, home_fireplace:home_insert), 
               names_to = "var", values_to = "value") %>% 
  group_by(home, var) %>% 
  arrange(home, var, winter_id) %>% 
  mutate(data_check = if_else(value == lead(value), 1, 0),
         data_check = if_else(is.na(data_check), lag(data_check), data_check)) %>% 
  filter(data_check == 0) %>% 
  select(area, home, winter_id, var, value) %>% 
  mutate(var2 = var,
         value2 = value) %>% 
  pivot_wider(names_from = "var2", values_from = "value2") %>% 
  ungroup() %>% 
  mutate_at(c("home_insert", "stove_age", "home_furnace", "home_woodstove", 
              "stove_cert", "home_type", "home_fireplace"), as.character) %>% 
  mutate_at(c("home_insert", "stove_age", "home_furnace", "home_woodstove", 
              "stove_cert", "home_type", "home_fireplace"), as.factor) %>% 
  mutate_at(c("home_insert", "stove_age", "home_furnace", "home_woodstove", 
              "stove_cert", "home_type", "home_fireplace"), as.numeric)

home_char_check_fact$value_num <- 
  coalesce(home_char_check_fact$home_insert, 
           home_char_check_fact$stove_age, 
           home_char_check_fact$home_furnace, 
           home_char_check_fact$home_woodstove, 
           home_char_check_fact$stove_cert, 
           home_char_check_fact$home_type, 
           home_char_check_fact$home_fireplace)

home_char_check_fact2 <- home_char_check_fact %>% 
  select(area, home, winter_id, var, value, value_num) %>% 
  group_by(home, var) %>% 
  mutate(data_check = if_else(abs(value_num - lead(value_num)) > 1, 1, 0),
         data_check = if_else(is.na(data_check), lag(data_check), data_check),
         value = as.character(value))

write_csv(home_char_check_fact2, paste0(file_path, "Output/home_char_check_fact.csv"))

home_char_check <- rbind(home_char_check_num, home_char_check_fact2) %>% 
  arrange(home, var, winter_id) %>% 
  mutate(data_check = if_else(is.na(value_num), 1, data_check))

write_csv(home_char_check, paste0(file_path, "Output/home_char_check.csv"))
```

### 2020-06-24 indoor_temp_rh_cleaning.Rmd
check home char across winters

---
title: "KidsAIR indoor temp and humidity cleaning"
author: "Ethan Walker"
date: "Started 13 Jan 2020, Updated 15 Jan 2020"
output: pdf_document
---

@@

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
```

# Load, format, and save indoor temp/rh data
```{r}
##### WMT #####
# Load data
temp_rh_initial_wmt <- read_csv("Input/WMT/wmt_indoor_temp_rh_20200113.csv")

wmt_temp_rh <- temp_rh_initial_wmt %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         rh_indoor_current = as.numeric(RelHumCur),
         rh_indoor_max = as.numeric(RelHumMax),
         rh_indoor_min = as.numeric(RelHumMin),
         temp_indoor_current = as.numeric(TempCCur),
         temp_indoor_max = as.numeric(TempCMax),
         temp_indoor_min = as.numeric(TempCMin),
         rh_temp_comments = as.character(`REPLACE(REPLACE(Comments`),
         area = "WMT") %>% 
  # replace -999 values with NA
  replace_with_na(replace = list(rh_indoor_current = -999, rh_indoor_max = -999,
                                 rh_indoor_min = -999, temp_indoor_current = -999,
                                 temp_indoor_max = -999, temp_indoor_min = -999)) %>% 
  replace_with_na(replace = list(rh_indoor_current = 0, rh_indoor_max = 0,
                                 rh_indoor_min = 0, temp_indoor_current = 0,
                                 temp_indoor_max = 0, temp_indoor_min = 0)) %>% 
  select(home_winter_id, area, rh_indoor_current, 
         rh_indoor_max, rh_indoor_min, temp_indoor_current, 
         temp_indoor_max, temp_indoor_min, rh_temp_comments) %>% 
  arrange(home_winter_id)

# Save as RDS 
# write_rds(wmt_temp_rh, "Output/wmt_temp_rh.rds")

##### NN #####
# Load data
temp_rh_initial_nn <- read_xlsx("Input/NN/nn_indoor_temp_rh_20200113.xlsx")

nn_temp_rh <- temp_rh_initial_nn %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         rh_indoor_current = as.numeric(RelHumCur),
         rh_indoor_max = as.numeric(RelHumMax),
         rh_indoor_min = as.numeric(RelHumMin),
         temp_indoor_current = as.numeric(TempCCur),
         temp_indoor_max = as.numeric(TempCMax),
         temp_indoor_min = as.numeric(TempCMin),
         rh_temp_comments = as.character(IgnoreReason),
         area = "NN") %>% 
  # replace 0 and -999 values with NA
  replace_with_na(replace = list(rh_indoor_current = -999, rh_indoor_max = -999,
                                 rh_indoor_min = -999, temp_indoor_current = -999,
                                 temp_indoor_max = -999, temp_indoor_min = -999)) %>% 
  replace_with_na(replace = list(rh_indoor_current = 0, rh_indoor_max = 0,
                                 rh_indoor_min = 0, temp_indoor_current = 0,
                                 temp_indoor_max = 0, temp_indoor_min = 0)) %>% 
  select(home_winter_id, area, rh_indoor_current, 
         rh_indoor_max, rh_indoor_min, temp_indoor_current, 
         temp_indoor_max, temp_indoor_min, rh_temp_comments) %>% 
  arrange(home_winter_id)

# Save as RDS 
# write_rds(nn_temp_rh, "Output/nn_temp_rh.rds")

##### AK #####
# Load data
temp_rh_initial_ak <- read_xlsx("Input/AK/ak_indoor_temp_rh_20200113.xlsx")

ak_temp_rh <- temp_rh_initial_ak %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         rh_indoor_current = as.numeric(RelHumCur),
         rh_indoor_max = as.numeric(RelHumMax),
         rh_indoor_min = as.numeric(RelHumMin),
         temp_indoor_current = as.numeric(TempCCur),
         temp_indoor_max = as.numeric(TempCMax),
         temp_indoor_min = as.numeric(TempCMin),
         rh_temp_comments = as.character(IgnoreReason),
         area = "AK") %>% 
  # replace -999 values with NA
  replace_with_na(replace = list(rh_indoor_current = -999, rh_indoor_max = -999,
                                 rh_indoor_min = -999, temp_indoor_current = -999,
                                 temp_indoor_max = -999, temp_indoor_min = -999)) %>% 
  replace_with_na(replace = list(rh_indoor_current = 0, rh_indoor_max = 0,
                                 rh_indoor_min = 0, temp_indoor_current = 0,
                                 temp_indoor_max = 0, temp_indoor_min = 0)) %>% 
  select(home_winter_id, area, rh_indoor_current, 
         rh_indoor_max, rh_indoor_min, temp_indoor_current, 
         temp_indoor_max, temp_indoor_min, rh_temp_comments) %>% 
  arrange(home_winter_id)

# Save as RDS 
# write_rds(ak_temp_rh, "Output/ak_temp_rh.rds")
```

# Join temp/rh data from different areas
```{r}
wmt_temp_rh <- read_rds("Output/wmt_temp_rh.rds")
nn_temp_rh <- read_rds("Output/nn_temp_rh.rds")
ak_temp_rh <- read_rds("Output/ak_temp_rh.rds")

@@

# write_rds(kids_indoor_temp_rh, "Output/kids_indoor_temp_rh.rds")
```

# Clean up all temp/rh data
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
kids_indoor_temp_rh <- read_rds(paste0(file_path, "Output/kids_indoor_temp_rh.rds"))
kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids.rds"))

indoor_temp_rh_clean <- kids_indoor_temp_rh %>% 
  left_join(kids_linked_ids, by = c("area", "home_winter_id"))

>>>>

---
title: "KidsAIR indoor temp and humidity cleaning"
author: "Ethan Walker"
date: "Started 13 Jan 2020, Updated 23 June 2020"
output: pdf_document
---

@@

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds")) 
```

# Load, format, and save indoor temp/rh data
```{r}
##### WMT #####
# Load data
temp_rh_initial_wmt <- read_csv(paste0(file_path, "Input/WMT/qtrack.csv"), 
                                  na = c("NULL", ""))

wmt_temp_rh <- temp_rh_initial_wmt %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         rh_indoor_current = as.numeric(RelHumCur),
         rh_indoor_max = as.numeric(RelHumMax),
         rh_indoor_min = as.numeric(RelHumMin),
         temp_indoor_current = as.numeric(TempCCur),
         temp_indoor_max = as.numeric(TempCMax),
         temp_indoor_min = as.numeric(TempCMin),
         area = "WMT") %>% 
  # replace -999 values with NA
  replace_with_na(replace = list(rh_indoor_current = -999, rh_indoor_max = -999,
                                 rh_indoor_min = -999, temp_indoor_current = -999,
                                 temp_indoor_max = -999, temp_indoor_min = -999)) %>% 
  replace_with_na(replace = list(rh_indoor_current = 0, rh_indoor_max = 0,
                                 rh_indoor_min = 0, temp_indoor_current = 0,
                                 temp_indoor_max = 0, temp_indoor_min = 0)) %>% 
  select(home_winter_id, area, rh_indoor_current, 
         rh_indoor_max, rh_indoor_min, temp_indoor_current, 
         temp_indoor_max, temp_indoor_min) %>% 
  arrange(home_winter_id)

# Save as RDS 
write_rds(wmt_temp_rh, paste0(file_path, "Output/wmt_temp_rh.rds"))

##### NN #####
# Load data
temp_rh_initial_nn <- read_csv(paste0(file_path, "Input/NN/qtrack.csv"), 
                                  na = c("NULL", ""))

nn_temp_rh <- temp_rh_initial_nn %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         rh_indoor_current = as.numeric(RelHumCur),
         rh_indoor_max = as.numeric(RelHumMax),
         rh_indoor_min = as.numeric(RelHumMin),
         temp_indoor_current = as.numeric(TempCCur),
         temp_indoor_max = as.numeric(TempCMax),
         temp_indoor_min = as.numeric(TempCMin),
         area = "NN") %>% 
  # replace 0 and -999 values with NA
  replace_with_na(replace = list(rh_indoor_current = -999, rh_indoor_max = -999,
                                 rh_indoor_min = -999, temp_indoor_current = -999,
                                 temp_indoor_max = -999, temp_indoor_min = -999)) %>% 
  replace_with_na(replace = list(rh_indoor_current = 0, rh_indoor_max = 0,
                                 rh_indoor_min = 0, temp_indoor_current = 0,
                                 temp_indoor_max = 0, temp_indoor_min = 0)) %>% 
  select(home_winter_id, area, rh_indoor_current, 
         rh_indoor_max, rh_indoor_min, temp_indoor_current, 
         temp_indoor_max, temp_indoor_min) %>% 
  arrange(home_winter_id)

# Save as RDS 
write_rds(nn_temp_rh, paste0(file_path, "Output/nn_temp_rh.rds"))

##### AK #####
# Load data
temp_rh_initial_ak <- read_csv(paste0(file_path, "Input/AK/qtrack.csv"), 
                                  na = c("NULL", ""))

ak_temp_rh <- temp_rh_initial_ak %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         rh_indoor_current = as.numeric(RelHumCur),
         rh_indoor_max = as.numeric(RelHumMax),
         rh_indoor_min = as.numeric(RelHumMin),
         temp_indoor_current = as.numeric(TempCCur),
         temp_indoor_max = as.numeric(TempCMax),
         temp_indoor_min = as.numeric(TempCMin),
         area = "AK") %>% 
  # replace -999 values with NA
  replace_with_na(replace = list(rh_indoor_current = -999, rh_indoor_max = -999,
                                 rh_indoor_min = -999, temp_indoor_current = -999,
                                 temp_indoor_max = -999, temp_indoor_min = -999)) %>% 
  replace_with_na(replace = list(rh_indoor_current = 0, rh_indoor_max = 0,
                                 rh_indoor_min = 0, temp_indoor_current = 0,
                                 temp_indoor_max = 0, temp_indoor_min = 0)) %>% 
  select(home_winter_id, area, rh_indoor_current, 
         rh_indoor_max, rh_indoor_min, temp_indoor_current, 
         temp_indoor_max, temp_indoor_min) %>% 
  arrange(home_winter_id)

# Save as RDS 
write_rds(ak_temp_rh, paste0(file_path, "Output/ak_temp_rh.rds"))
```

# Join temp/rh data from different areas
```{r}
wmt_temp_rh <- read_rds(paste0(file_path, "Output/wmt_temp_rh.rds"))
nn_temp_rh <- read_rds(paste0(file_path, "Output/nn_temp_rh.rds"))
ak_temp_rh <- read_rds(paste0(file_path, "Output/ak_temp_rh.rds"))

@@

write_rds(kids_indoor_temp_rh, paste0(file_path, "Output/kids_indoor_temp_rh.rds"))
```

# Clean up all temp/rh data
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
kids_indoor_temp_rh <- read_rds(paste0(file_path, "Output/kids_indoor_temp_rh.rds"))
kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds"))

indoor_temp_rh_clean <- kids_indoor_temp_rh %>% 
  mutate(area = as.factor(area)) %>% 
  right_join(kids_linked_ids, by = c("area", "home_winter_id"))

summary(indoor_temp_rh_clean)

### 2020-06-24 moisture_data_cleaning.Rmd
check home char across winters

---
title: "KidsAIR initial moisture content data work"
author: "Ethan Walker"
date: "Started 19 Nov 2019, Updated 14 Jan 2020"
output: pdf_document
---

@@

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
```

# Load, format, and save fuel moisture content data
```{r}
##### WMT #####
# Load data
moisture_initial_wmt <- read_xlsx("Input/WMT/wmt_kids_moisture_20191122.xlsx")

wmt_moisture <- moisture_initial_wmt %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         moisture_date = ymd(CompletedDate),
         moisture_1 = as.numeric(Moisture1),
         moisture_2 = as.numeric(Moisture2),
         moisture_3 = as.numeric(Moisture3),
         moisture_ave = as.numeric(MoistureAve),
         moisture_split = as.factor(Split),
         area = "WMT") %>% 
  # replace 0 and -999 values with NA
  replace_with_na(replace = list(moisture_1 = 0, moisture_2 = 0,
                                 moisture_3 = 0, moisture_ave = 0)) %>% 
  replace_with_na(replace = list(moisture_1 = -999, moisture_2 = -999,
                                 moisture_3 = -999)) %>% 
  # gather 3 moisture samples to calculate average
  gather("sample", "value", c("moisture_1":"moisture_3")) %>% 
  group_by(home_winter_id, moisture_date) %>% 
  mutate(moisture_ave = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # spread 3 moisture samples back to wide format
  spread(sample, value) %>% 
  select(home_winter_id, area, moisture_date, 
         moisture_1, moisture_2, moisture_3, 
         moisture_ave, moisture_split) %>% 
  arrange(home_winter_id, moisture_date)

# Save as RDS 
# write_rds(wmt_moisture, "Output/wmt_moisture.rds")

##### NN #####
# Load data
moisture_initial_nn <- read_xlsx("Input/NN/nn_moisture_20200113.xlsx")

nn_moisture <- moisture_initial_nn %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         moisture_date = ymd(CompletedDate),
         moisture_1 = as.numeric(Moisture1),
         moisture_2 = as.numeric(Moisture2),
         moisture_3 = as.numeric(Moisture3),
         moisture_ave = as.numeric(MoistureAve),
         moisture_split = as.factor(Split),
         area = "NN") %>% 
  # replace 0 and -999 values with NA
  replace_with_na(replace = list(moisture_1 = 0, moisture_2 = 0,
                                 moisture_3 = 0, moisture_ave = 0)) %>% 
  replace_with_na(replace = list(moisture_1 = -999, moisture_2 = -999,
                                 moisture_3 = -999)) %>% 
  # gather 3 moisture samples to calculate average
  gather("sample", "value", c("moisture_1":"moisture_3")) %>% 
  group_by(home_winter_id, moisture_date) %>% 
  mutate(moisture_ave = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # spread 3 moisture samples back to wide format
  spread(sample, value) %>% 
  select(home_winter_id, area, moisture_date, 
         moisture_1, moisture_2, moisture_3, 
         moisture_ave, moisture_split) %>% 
  arrange(home_winter_id, moisture_date)

# Save as RDS 
# write_rds(nn_moisture, "Output/nn_moisture.rds")

##### AK #####
# Load data
moisture_initial_ak <- read_xlsx("Input/AK/ak_moisture_20200113.xlsx")

ak_moisture <- moisture_initial_ak %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         moisture_date = ymd(CompletedDate),
         moisture_1 = as.numeric(Moisture1),
         moisture_2 = as.numeric(Moisture2),
         moisture_3 = as.numeric(Moisture3),
         moisture_ave = as.numeric(MoistureAve),
         moisture_split = as.factor(Split),
         area = "AK") %>% 
  # replace 0 and -999 values with NA
  replace_with_na(replace = list(moisture_1 = 0, moisture_2 = 0,
                                 moisture_3 = 0, moisture_ave = 0)) %>% 
  replace_with_na(replace = list(moisture_1 = -999, moisture_2 = -999,
                                 moisture_3 = -999)) %>% 
  # gather 3 moisture samples to calculate average
  gather("sample", "value", c("moisture_1":"moisture_3")) %>% 
  group_by(home_winter_id, moisture_date) %>% 
  mutate(moisture_ave = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # spread 3 moisture samples back to wide format
  spread(sample, value) %>% 
  select(home_winter_id, area, moisture_date, 
         moisture_1, moisture_2, moisture_3, 
         moisture_ave, moisture_split) %>% 
  arrange(home_winter_id, moisture_date)

# Save as RDS 
# write_rds(ak_moisture, "Output/ak_moisture.rds")
```

# Join moisture data from different areas
```{r}
wmt_moisture <- read_rds("Output/wmt_moisture.rds")
nn_moisture <- read_rds("Output/nn_moisture.rds")
ak_moisture <- read_rds("Output/ak_moisture.rds")

@@

# write_rds(kids_moisture, "Output/kids_moisture.rds")
```

# Clean up all moisture data
```{r}
kids_moisture <- read_rds("Output/kids_moisture.rds")
kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")

moisture_clean <- kids_moisture %>% 
  left_join(kids_linked_ids, by = c("area", "home_winter_id"))

# write_rds(moisture_clean, "Output/moisture_clean.rds")
```

>>>>

---
title: "KidsAIR initial moisture content data work"
author: "Ethan Walker"
date: "Started 19 Nov 2019, Updated 23 June 2020"
output: pdf_document
---

@@

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds")) 
```

# Load, format, and save fuel moisture content data
```{r}
##### WMT #####
# Load data
moisture_initial_wmt <- read_csv(paste0(file_path, "Input/WMT/moisture.csv"), 
                                  na = c("NULL", ""))

wmt_moisture <- moisture_initial_wmt %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         moisture_date = ymd(CompletedDate),
         moisture_1 = as.numeric(Moisture1),
         moisture_2 = as.numeric(Moisture2),
         moisture_3 = as.numeric(Moisture3),
         moisture_ave = as.numeric(MoistureAve),
         moisture_split = as.factor(Split),
         area = "WMT") %>% 
  # replace 0 and -999 values with NA
  replace_with_na(replace = list(moisture_1 = 0, moisture_2 = 0,
                                 moisture_3 = 0, moisture_ave = 0)) %>% 
  replace_with_na(replace = list(moisture_1 = -999, moisture_2 = -999,
                                 moisture_3 = -999)) %>% 
  # gather 3 moisture samples to calculate average
  gather("sample", "value", c("moisture_1":"moisture_3")) %>% 
  group_by(home_winter_id, moisture_date) %>% 
  mutate(moisture_ave = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # spread 3 moisture samples back to wide format
  spread(sample, value) %>% 
  select(home_winter_id, area, moisture_date, 
         moisture_1, moisture_2, moisture_3, 
         moisture_ave, moisture_split) %>% 
  replace_with_na(replace = list(moisture_ave = "NaN")) %>% 
  arrange(home_winter_id, moisture_date)

# Save as RDS 
write_rds(wmt_moisture, paste0(file_path, "Output/wmt_moisture.rds"))

##### NN #####
# Load data
moisture_initial_nn <- read_csv(paste0(file_path, "Input/NN/moisture.csv"), 
                                  na = c("NULL", ""))

nn_moisture <- moisture_initial_nn %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         moisture_date = ymd(CompletedDate),
         moisture_1 = as.numeric(Moisture1),
         moisture_2 = as.numeric(Moisture2),
         moisture_3 = as.numeric(Moisture3),
         moisture_ave = as.numeric(MoistureAve),
         moisture_split = as.factor(Split),
         area = "NN") %>% 
  # replace 0 and -999 values with NA
  replace_with_na(replace = list(moisture_1 = 0, moisture_2 = 0,
                                 moisture_3 = 0, moisture_ave = 0)) %>% 
  replace_with_na(replace = list(moisture_1 = -999, moisture_2 = -999,
                                 moisture_3 = -999)) %>% 
  # gather 3 moisture samples to calculate average
  gather("sample", "value", c("moisture_1":"moisture_3")) %>% 
  group_by(home_winter_id, moisture_date) %>% 
  mutate(moisture_ave = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # spread 3 moisture samples back to wide format
  spread(sample, value) %>% 
  select(home_winter_id, area, moisture_date, 
         moisture_1, moisture_2, moisture_3, 
         moisture_ave, moisture_split) %>% 
  replace_with_na(replace = list(moisture_ave = "NaN")) %>% 
  arrange(home_winter_id, moisture_date)

# Save as RDS 
write_rds(nn_moisture, paste0(file_path, "Output/nn_moisture.rds"))

##### AK #####
# Load data
moisture_initial_ak <- read_csv(paste0(file_path, "Input/AK/moisture.csv"), 
                                  na = c("NULL", ""))

ak_moisture <- moisture_initial_ak %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         moisture_date = ymd(CompletedDate),
         moisture_1 = as.numeric(Moisture1),
         moisture_2 = as.numeric(Moisture2),
         moisture_3 = as.numeric(Moisture3),
         moisture_ave = as.numeric(MoistureAve),
         moisture_split = as.factor(Split),
         area = "AK") %>% 
  # replace 0 and -999 values with NA
  replace_with_na(replace = list(moisture_1 = 0, moisture_2 = 0,
                                 moisture_3 = 0, moisture_ave = 0)) %>% 
  replace_with_na(replace = list(moisture_1 = -999, moisture_2 = -999,
                                 moisture_3 = -999)) %>% 
  # gather 3 moisture samples to calculate average
  gather("sample", "value", c("moisture_1":"moisture_3")) %>% 
  group_by(home_winter_id, moisture_date) %>% 
  mutate(moisture_ave = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # spread 3 moisture samples back to wide format
  spread(sample, value) %>% 
  select(home_winter_id, area, moisture_date, 
         moisture_1, moisture_2, moisture_3, 
         moisture_ave, moisture_split) %>% 
  replace_with_na(replace = list(moisture_ave = "NaN")) %>% 
  arrange(home_winter_id, moisture_date)

# Save as RDS 
write_rds(ak_moisture, paste0(file_path, "Output/ak_moisture.rds"))
```

# Join moisture data from different areas
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

wmt_moisture <- read_rds(paste0(file_path, "Output/wmt_moisture.rds"))
nn_moisture <- read_rds(paste0(file_path, "Output/nn_moisture.rds"))
ak_moisture <- read_rds(paste0(file_path, "Output/ak_moisture.rds"))

@@

write_rds(kids_moisture, paste0(file_path, "Output/kids_moisture.rds"))
```

# Clean up all moisture data
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
kids_moisture <- read_rds(paste0(file_path, "Output/kids_moisture.rds"))
kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds"))

moisture_clean <- kids_moisture %>% 
  mutate(area = as.factor(area)) %>% 
  right_join(kids_linked_ids, by = c("area", "home_winter_id")) %>% 
  # lower range on instrument = 5%
  # therefore, changing values < 5 to 5
  mutate(moisture_1 = if_else(moisture_1 < 5, 5, moisture_1),
         moisture_2 = if_else(moisture_2 < 5, 5, moisture_2),
         moisture_3 = if_else(moisture_3 < 5, 5, moisture_3),
         moisture_ave = if_else(moisture_ave < 5, 5, moisture_ave))

summary(moisture_clean)

write_rds(moisture_clean, paste0(file_path, "Output/moisture_clean.rds"))
```
